%! program = pdflatexmk

% VoteSecure FAQ
% Copyright (C) 2025 Free & Fair

\documentclass[12pt,letter]{article}

%%% PACKAGES
\usepackage[in]{fullpage}
\usepackage{tabularray}
\usepackage[pdftex,bookmarks,colorlinks,breaklinks]{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\setlength {\marginparwidth}{2cm}
\usepackage{todonotes}
\usepackage{enumitem}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}
\parindent 0pt
\textheight 9in
\topmargin 0in
\setlength{\parskip}{10pt plus 1pt minus 1pt}

\usepackage{tocloft}
\setlength\cftparskip{2pt}

% format FAQ date in British English, but without the ordinal suffix
\usepackage[en-GB]{datetime2}
\DTMlangsetup[en-GB]{ord=omit}

\renewcommand{\subsectionautorefname}{section}
\renewcommand{\subsubsectionautorefname}{section}

% by default, don't put headings on tables
\DefTblrTemplate{firsthead, middlehead,lasthead}{default}{}

\pagestyle{empty}

%%% BEGIN DOCUMENT
\begin{document}

\begin{center}
{\Large \textbf{VoteSecure FAQ}}
\end{center}
\vspace{-6pt}

\textit{Free \& Fair} \hspace{\fill} \textit{\DTMtoday}

\vspace{-12pt}

\rule{\textwidth}{1pt}

The concept of mobile voting naturally raises many questions, both about the idea itself and about the role of the VoteSecure library in the implementation of any publicly-available mobile voting system. This document contains our answers to questions we have been asked, and will be updated with future questions and answers as they arise.

% Q & A go here, we can format them in a consistent manner later on

\section{General Project Questions}

\begin{itemize}

    \item \emph{Q:} Where should I start reading?

    \emph{A:} It's best to start at the project page on our website and go from there. \\
    \href{https://freeandfair.us/votesecure/}{https://freeandfair.us/votesecure/}

    \item \emph{Q:} Is there more information coming out soon?

    \emph{A:} Every time we merge new content into the \texttt{main} branch of the project it is mirrored to GitHub.  So, while we have a few active branches of work (e.g., on mechanizing the protocol in Tamarin) that will not be public until they are merged into \texttt{main}, the vast majority of the work is public already, and it will all be public when it is complete.

    \item \emph{Q:} Is most or all of the technology developed by the Danish company Assembly Voting gone? Are you making something different?

    \emph{A:} Yes, all of Assembly Voting's work is gone.  We analyzed their cryptographic protocol and implementation; \href{https://github.com/FreeAndFair/Transparency/blob/master/Project_Artifacts/Tusk_2024_Review/cryptographic-protocol-review-and-gap-analysis.pdf}{our report summarizing that analysis} is publicly available in our \href{https://github.com/FreeAndFair/Transparency}{Transparency repository} on GitHub.

    When this project started we had originally thought that portions of that protocol might be reusable, but as we performed a more detailed security analysis of the protocol we found so many flaws that it became clear we had to start from scratch.

    \item \emph{Q:} Will white hat hackers be given legal protection and a reasonable lead time prior to the system being used in any election?  If so, will there be any publication, NDA, or other constraints on what the hackers may say/write?

    \emph{A:} We are only developing the core cryptographic library, and we cannot speak for system implementations; however, we welcome security analysis of both the cryptographic protocol and (when available) its implementation. The project repository contains \href{https://github.com/FreeAndFair/MobileVotingCoreCryptography/blob/main/SECURITY.md}{instructions for reporting security issues} that follow standard \href{https://en.wikipedia.org/wiki/Coordinated_vulnerability_disclosure}{coordinated vulnerability disclosure (CVD)} guidelines. Other than requesting that they follow CVD guidelines, we will not place any restrictions on what anybody can write or say.

\end{itemize}


\section{Cryptography}

\begin{itemize}

    \item \emph{Q:} Are you using a mixnet?

    \emph{A:} We are using a mixnet in the first version, because that's the most straightforward way to produce CVRs that can be used to generate normal paper ballots. The election officials (EOs) Tusk has been working extensively with want the back end of the system to either be a human-centric remarking workflow exactly like they use today, or to produce PDFs of normal paper ballots that are readable by their existing scanners.  Note that we are only implementing the cryptographic protocol, not this back end, in our current work.

    \item \emph{Q:} Do you support homomorphic tallying?

    \emph{A:} While a homomorphic tally is possible for our protocol, and is a variant we have examined closely, we are not implementing it at this time.  Such a back end has an entirely different workflow for EOs---one that they are not familiar or comfortable with as of now, as we understand it. It also presents issues with respect to certain ballot features, such as write-in candidates.

    \item \emph{Q:} Do you use a Benaloh-like challenge as a part of the voter's user experience as seen in the original AV protocol?

    \emph{A:} As you will see in our analysis of the original protocol, the AV Benaloh-like challenge had fundamental security flaws, which perhaps contributed to the positive outcomes in their usability studies.  We have had to think hard about the voter experience in order to obtain a user experience that fulfills E2E-V requirements and yet is usable by all voters.

\end{itemize}

\section{Election Officials}

\begin{itemize}

    \item \emph{Q:} Aren't there real concerns about the security (the ordinary kind) of the IT systems to facilitate elections, and the feasibility of ordinary EOs doing the key management and key usage?

    \emph{A:} While key management is a necessary part of most systems that perform cryptography, we hold low concern about EOs ability to generate, manage, and use keys in our system.  The protocol itself identifies key management problems and threats as a part of its overall CONOPS and threat model, as mistakes and adversaries focused on key material are important. However, given modern key generation and management systems that are now widely deployed in public, including key managers in OSs, browsers, and password managers, as well as more esoteric, secure, and somewhat less user-friendly devices like Yubikeys and Google Titan keys, we are in a much better place today than we were just a few years ago.

    \item \emph{Q:} I liked (in theory) the basic idea of a digital process very similar to the paper absentee ballot process: the ballot is cloaked and accompanied by an affidavit; and it is the EO's responsibility to decide (based on the affidavit and other stuff) whether each ballot is admissible, and only if so is the cloaked ballot queued up for counting. I don't know whether the digital version of that approach (including the same literal affidavit document as in paper AB) can be done in the scheme you're using, but I am curious if it is a goal, and if so whether it can be achieved using SW that is usable by ordinary EOs, and deployed on IT systems that can be protected by EO IT staff and systems.

    \emph{A:} The goal is indeed to support that kind of workflow as a part of our protocol.  It is what EOs want, and they are the domain experts.  Thus, in our protocol the authentication of voters is delegated to a black box component with a generic interface, set of behaviors, and risk characterization.

\end{itemize}

\section{Security}

\begin{itemize}

    \item \emph{Q:} How will the voter be authenticated?

    \emph{A:} Authentication is a ``black box'' in our protocol, meaning that system implementers and EOs must decide upon appropriate authentication processes/providers for their individual implementations and applications. We expect that identity verification systems like \href{https://clearme.com/}{CLEAR} will be used for this purpose, though other approaches (e.g., a digital equivalent to the signed affidavits that accompany mail-in ballots today) can be implemented as well.

    \item \emph{Q:} How will the system defend against distributed denial of service (DDoS) attacks?

    \emph{A:} We are only developing the core cryptographic library, and we cannot speak for system implementations; however, we expect that system implementers will use industry-standard methods for defending Internet-facing systems against DDoS attacks (i.e., \href{https://cloudflare.com/}{Cloudflare} and similar services).

    \item \emph{Q:} How can underfunded and under-resourced local election officials protect their systems from attacks, both external and from insiders? What about ransomware?

    \emph{A:} As \href{https://github.com/FreeAndFair/MobileVotingCoreCryptography/releases/download/latest/threat-model.pdf}{our threat model} states, we are explicitly not addressing these sorts of system security issues as part of this project; these will need to be addressed by system implementers. However, we have designed the cryptographic protocol such that (1) voter privacy cannot be compromised without using private key material belonging to a sufficiently large number of election trustees, and (2) ballots cannot be changed in undetectable ways after being submitted by voters.

    \item \emph{Q:} Can voters verify their ballots using a separate device from the one they used to vote?

    \emph{A:} Yes. The ballot check application can be run from any device---voters might choose to run it from a web browser or separate app on the same device they used to vote, but verification should ideally take place on an independent device (or more than one such device) to provide stronger protection against a compromised voting application. This is a key security feature that distinguishes VoteSecure from some other E2E-V systems where verification typically occurs on the same device.

\end{itemize}

\section{Usability}

\begin{itemize}

    \item \emph{Q:} One issue with E2E-V is that it's not easy for non-technical people to understand why is works. How do you propose to address the trust issue?

    \emph{A:} We have attempted to make the protocol's implementation of end-to-end verifiability straightforward; however, work will clearly need to be done to present it to voters in an approachable, understandable way. This will involve user interface design work for any voting and ballot check applications that implement the protocol, as well as other explanatory resources (documents, presentations, etc.).

\end{itemize}

\section{Comparison with Other E2E-V Systems}

\begin{itemize}

    \item \emph{Q:} How does VoteSecure compare with other end-to-end verifiable voting systems like Helios, Belenios, and the Swiss Post system?

    \emph{A:} VoteSecure shares some architectural similarities with other E2E-V systems but has several important differences in its threat model, privacy guarantees, and verification approach. A detailed technical comparison is available in the \href{https://github.com/FreeAndFair/MobileVotingCoreCryptography/blob/main/docs/voting-systems-comparison.md}{Voting Systems Comparison} document. Here is a summary:

    \textbf{Helios:} Like Helios, VoteSecure uses ElGamal encryption and provides cast-as-intended, recorded-as-cast, and tallied-as-cast verification. However, VoteSecure differs in several key ways: (1) VoteSecure uses a mixnet for decryption (in the first version) rather than homomorphic tallying, which better supports traditional election official workflows and handles write-in candidates; (2) VoteSecure's ballot check application can run on a separate device from the voting application, providing stronger cast-as-intended verification against compromised voting clients; (3) VoteSecure is designed for mobile voting scenarios with explicit threat modeling of malicious applications on the voter's device.

    \textbf{Belenios:} Belenios is an open-source E2E-V system that, like VoteSecure, emphasizes formal verification of its cryptographic protocols. Both systems provide individual and universal verifiability. VoteSecure's approach differs in that it uses rigorous digital engineering (RDE) methodology throughout, with formal refinement from high-level models down to implementation. VoteSecure also explicitly addresses the threat of advanced persistent threats (APTs) across all system environments, including mobile devices, which is particularly important for Internet voting scenarios.

    \textbf{Swiss Post System:} The Swiss Post e-voting system (now discontinued after security issues were discovered) used a different approach with individual verification codes. VoteSecure's verification approach is more transparent and allows for independent verification by third parties through the public bulletin board. Unlike Swiss Post, VoteSecure is fully open source, enabling unrestricted security analysis by researchers and the public.

    \textbf{Receipt-freeness:} Like Helios and most practical E2E-V systems designed for remote voting, VoteSecure does not provide receipt-freeness. A minor client-side modification could retain the randomness used in encryption, allowing voters to prove how they voted. This is an inherent trade-off in remote voting systems where the voting environment cannot be controlled---providing strong verifiability while preventing vote buying and coercion is cryptographically challenging. The threat of coercion is acknowledged in our threat model.

    \textbf{Trust Model:} VoteSecure's trust model assumes that privacy requires a threshold of trustees to maintain secrecy of their key shares, similar to other mixnet-based systems. For integrity, the system is designed so that ballots cannot be changed undetectably after submission, even if some trustees are malicious. The public bulletin board and cryptographic proofs enable universal verifiability, meaning anyone can verify the correctness of the tally.

\end{itemize}

\end{document}
